Open Software - Restricted Data: A case study.

### 2016-01-23 report1_high_level/manuscript.Rmd
add different image

Draft: \today

@@

\vspace{2mm}\hrule
\paragraph*{Background:} This essay was written to accompany the material presented as a
speedtalk and poster at the National Climate Change Adaptation
Research Facility Conference 'Climate Adaptation knowledge and
partnership', June 2013, Sydney.  The poster and slideshow are both available to download from this website: [http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/presentations-nccarf-2013/](http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/presentations-nccarf-2013/). The latest version of this working paper is available to download from the website: [http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/report1_high_level/manuscript.pdf](http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/report1_high_level/manuscript.pdf).

\paragraph*{Methods:} The paper reports on a project to build tools and procedures for enhancing open and transparent analysis of restricted datasets.
Some datasets such as suicide or climate
change scenarios need to be accessed in a restricted way.  On the
other hand scientists need to make their methods, models and
assumptions transparent and available for scientific debate even
though the datasets may require authorisation to access.

@@

```{r, eval = F, echo = F}
dir()
dir("~/data")
indir <- "~/data/GARNAUT mental health archive/1 Data/Rain"
dir(indir, pattern = "jpg")
filelist  <- c("A1BDRY_RainSD07.jpg",             "A1BWET_RainSD07.jpg"            )
for(fi in filelist){
  #file.copy(file.path(indir, fi), fi)
  txt <- sprintf("\\begin{figure}[!h]
\\centering
\\includegraphics[width=\\textwidth]{%s}
\\caption{%s}
\\label{fig:%s}
\\end{figure}
", fi, fi, fi)
  cat(txt)
}
```

@@

\clearpage

Below are pictorial represenations of the climate change scenarios estimated annual total rainfall for each Statistical Division (SD) census region across the country.  SDs are used because this is the geographical unit at which the suicide rates were analysed.

\begin{figure}[!h]
\centering
\includegraphics[width=.5\textwidth]{A1BDRY_RainSD07.jpg}
\caption{A1BDRY RainSD07}
\label{fig:A1BDRYRainSD07}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=.5\textwidth]{A1BWET_RainSD07.jpg}
\caption{A1BWET RainSD07}
\label{fig:A1BWET-RainSD07}
\end{figure}

@@

The results of our project are applicable more generally than just
reproducible research. For instance in relation to issues of
governance and management at the multidisciplinary institutions and
multi-institutional projects it is vitally important that data sharing
is enabled, while some data are still restricted because access requires
approval and authorisation, or because data providers wish to be informed of re-use and identifying the profile of
users of the data.  It is important that data access
can be made restrictive WITHOUT impeding the progress of the local
science agenda (collaborations, workshops, papers etc) and keeping the
relevant data custodian parties informed about what is happening with
the release of their datasets.  The open-ness of the analytical
software also has a positive effect on the value of the data
infrastructure (through education and outreach activities) without
risking any unethical or negligent use of these datasets.

@@

In summary we found that:

- Climate data can be attached to suicide and census data in a secure computing environment.
- Future drought associated deaths can be estimated.
- These estimates are very uncertain, contentious and difficult to justify.
- Data management and analysis technology such as that presented is needed to enable rigorous and transparent exploration that challenges or verifies published results.

This system described:

- Enables data analysis in a safe environment.
- Allows comparison of multiple climate scenarios and assumptions.
- Demonstrated with a climate/health impact assessment.
- Addresses the reproducibility crisis.

>>>>

@@

Unpublished working paper\\
Draft: \today

\vspace{2mm}\hrule
\paragraph*{Background:} This unpublished working paper was written to accompany the material presented as a
speedtalk and poster at the National Climate Change Adaptation
Research Facility Conference 'Climate Adaptation knowledge and
partnership', June 2013, Sydney.  The poster and slideshow are both available to download from this website: [http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/presentations-nccarf-2013/](http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/presentations-nccarf-2013/). The latest version of this working paper is available to download from the website: [http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/report1_high_level/manuscript.pdf](http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/report1_high_level/manuscript.pdf).

\paragraph*{Methods:} We report on a project to build tools and procedures for enhancing open and transparent analysis of restricted datasets.
Some datasets such as suicide or climate
change scenarios need to be accessed in a restricted way.  On the
other hand scientists need to make their methods, models and
assumptions transparent and available for scientific debate even
though the datasets may require authorisation to access.

@@

@@

\clearpage
Below are pictorial represenations of the climate change scenarios estimated seasonal rainfall for each Statistical Division (SD) census region across the country.  SDs are used because this is the geographical unit at which the suicide rates were analysed.

```{r, eval = F, echo = F}
dir()
dir("~/data")
indir <- "~/data/GARNAUT mental health archive/1 Data/Rain/seasonal/"
dir(indir, pattern = "jpg")
fi  <- "SeasonalA1FIR1andR2_final.jpg"
cap <- "Seasonal Rainfall under alternate future predicted scenarios (A1FI) with wet or dry assumptions"
#file.copy(file.path(indir, fi), fi)
txt <- sprintf("\\begin{figure}[!h]
\\centering
\\includegraphics[width=\\textwidth]{%s}
\\caption{%s}
\\label{fig:%s}
\\end{figure}
", fi, cap, fi)
  cat(txt)

\begin{figure}[!h]
\centering
\includegraphics[width=.78\textwidth]{SeasonalA1FIR1andR2_final.jpg}
\caption{Seasonal rainfall under alternate future predicted scenarios (A1FI) with wet or dry assumptions}
\label{fig:SeasonalA1FIR1andR2_final.jpg}
\end{figure}
\clearpage

@@

The results of our project are applicable more generally than just
reproducible research. For instance in relation to issues of
governance and management at the multidisciplinary institutions and
multi-institutional projects it is vitally important that data sharing
is enabled, while some data are still restricted because access requires
approval and authorisation, or because data providers wish to be informed of re-use and identifying the profile of
users of the data.  It is important that data access
can be made restrictive without impeding the progress of the local
science agenda (collaborations, workshops, papers etc) and keeping the
relevant data custodian parties informed about what is happening with
the release of their datasets.  The open-ness of the analytical
software also has a positive effect on the value of the data
infrastructure (through education and outreach activities) without
risking any unethical or negligent use of these datasets.

@@

In summary we found that the historical weather and climate change data can be attached to suicide and census data in a secure computing environment. The future drought associated deaths can then be estimated using statistical tools available in an online virtual machine.  This facilitates collaboration because the data can be accessed remotely from collaborator\'s offices (or even workshops and conference venues) while adhering to requirements for secure data storage and access constraints.

In the case study presented it should be noted that these estimates are very uncertain, contentious and difficult to justify. However, this reinforces the argument that data management and analysis technology such as those presented are needed to enable rigorous and transparent exploration of methods and assumptions. In this way a community of like-minded scientists can challenge or verify published results.

In conclusion the system described enables data analysis in a safe environment;
allows comparison of multiple climate scenarios and assumptions;
is demonstrated with a climate/health impact assessment and addresses the 'reproducibility crisis'.
